Solutions to Integer Programming from Quantum Annealing

### 2020-05-11 paper/main.tex
Formatting
First reordering

\documentclass[prd,twocolumn,tightenlines,preprintnumbers,showpacs,superscriptaddress,notitlepage,nofootinbib,eqsecnum,floatfix,longbibliography]{revtex4}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{bm,bbm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{dsfont}

@@

\section{INTRODUCTION}
\label{sec:introduction}

@@

Formally, ILP is defined as the optimization of a linear function
{\color{blue}
\begin{align}
f(x) = &\min\limits_{x}(\sum_i c_i x_i),\\
\textrm{subject to} \quad & \sum_i A_{ai}x_i +b_a \leq 0 \\
& x_i \geq 0,\\
& x_i \in \mathbb{Z}
\end{align}
}
where $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations. The inequality constraints can be rewritten as an equality by introducing slack variables $s$ such that the ILP becomes
{\color{blue}
\begin{align}
f(x) = &\min(\sum_i c_i x_i),\\
\textrm{subject to} \quad & \sum_i A_{a i}x_i + s_a + b_a = 0,\\
& s_a \geq 0,\\
& s_a \in \mathbb{Z},\\
& x_i \geq 0,\\
& x_i \in \mathbb{Z}.
\end{align}
}
Since quantum annealing solves the QUBO problem, we extend the proposed algorithm to solve integer quadratic optimization problems such that
{\color{blue}
\begin{align}
f(x) = \min\limits_{x}(\sum_{ij} x_i d_{ij} x_j + \sum_i c_i x_i)
\end{align}
}
without the introduction of auxiliary qubits.

@@

\section{RESULTS}
\label{sec:results}
We map integer variables $z_i$ to qubits under the following transformation~\cite{Chang:2018uoc}
\begin{align}
z_i = & \sum_{r=0}^{R_i-1} 2^r \psi_{ri}
\label{eq:int_to_bin}
\end{align}
where $\psi_{ri} \in \{0, 1\}$ while the number of qubits used to represent the $i$-th integer variable is allowed to vary with $R_i$. The transformation can in general be express as a rectangular matrix. For example, a vector of two integer variables $z_0$ and $z_1$ represented by one and two qubits respectively is given as
\begin{align}
\begin{pmatrix}
z_0\\
z_1
\end{pmatrix}
= &
\begin{pmatrix}
2^0 & 0 & 0\\
0 & 2^0 & 2^1
\end{pmatrix}
\begin{pmatrix}
\psi_{00}\\
\psi_{01}\\
\psi_{11}
\end{pmatrix}
\equiv T^z \begin{pmatrix}
\psi_{00}\\
\psi_{01}\\
\psi_{11}
\end{pmatrix}
\end{align}
The transformation can be reduced to a tensor product if $R_i$ is a constant for all $i$ such that
\begin{align}
\mathcal{R} = & \begin{pmatrix} 2^0 & \dots & 2^{R-1}\end{pmatrix},\\
\mathcal{Z} = & \begin{pmatrix} z_0 & \dots & z_{N-1}\end{pmatrix},\\
|\mathds{1}| = & |\mathcal{Z}|,\\
T^z = & \mathds{1}\otimes \mathcal{R}.
\end{align}

@@

\subsection{Integer Programming Mapping to QUBO}
\label{sec:results:ilp}
The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
\chi^2 = & \left[\Psi^x_{\mu} T^x_{\mu i}d_{ij} T^x_{j \nu}\Psi^x_\nu + c_i T^x_{i\mu} \Psi^x_\mu \right. \nonumber\\
&\left.+ p (A_{a b} T^x_{b \mu} \Psi^x_{\mu} + T^s_{a \alpha} \Psi^s_\alpha + b_a)^2 \right],\\
\Psi = & \begin{pmatrix}
\Psi^x\\
\Psi^s
\end{pmatrix},
\end{align}
where $p$ is the strength of the penalty factor which must be set large enough such that the constraints are satisfied under minimization. The objective function can be represented as a QUBO Hamiltonian
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
Q_{xx} & Q_{xs}\\
Q_{sx} & Q_{ss} 
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + pb^2,\\
\equiv & \Psi^T Q \Psi + C,
\label{eq:matrix_form}
\end{align}
where
\begin{align}
Q_{xx} = & {T^{x}}^T \left[ d + p A^T A + p \mathrm{Diag}_{x} \left(A^T b + b^T A\right) \right] T^x \nonumber \\
&+ \mathrm{Diag}_{\psi^x}(c) T^x,\\
Q_{xs} = & Q_{sx}^T = p {T^{x}}^T A^T T^s,\\
Q_{ss} = & p\left[ {T^{s}}^T T^s + \mathrm{Diag}_{\psi^s}\left( {T^{s}}^T b + b^T T^s\right) \right].
\end{align}
The function $\mathrm{Diag}_{z}(v)$ embeds a vector $v$ into a diagonal matrix in subspace $z$, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

\subsection{Application to the Dominating Set Problem}
\label{sec:results:mds}

@@

\subsubsection{Minimum Dominating Set}
The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
f(x) = &\min\left(\sum_{\nu \in V} \psi^x_{\nu}\right),\\
\textrm{subject to} \quad & \psi^x_{\nu} + \sum_{\mu \in \mathit{N}(\nu)} \psi^x_{\mu} \geq 1,\\
& \psi^x_{\nu} \in \{0, 1\}
\end{align}
where the dimension of the dependent variable $x$ is the number of vertices $\overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\psi^x$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(\nu)$ is in $D$. For each vertex in $V$ we introduce slack variables $s_{\nu} = T^s_{\nu \alpha} \psi^s_{\alpha}$ to encode the inequality constraint such that
\begin{align}
f(x) = &\min(\sum_{\nu\in V} \psi^x_{\nu}),\\
\textrm{subject to} \quad & \psi^x_{\nu} + J_{\nu \mu} \psi^x_{\mu}- T^s_{\nu \alpha} \psi^s_{\alpha}  - 1 = 0,\\
& \psi_{\nu} \in \{0, 1\},\\
& 0 \leq s_{\nu} \leq |\mathrm{N}(\nu)|\\
& s_{\nu} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$ (symmetric and zero diagonal), the adjacency matrix for $G$.
The algorithm uses $N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits---$\overline{\overline{V}}$ qubits to encode the vertices and another $\sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits to embed the slack variables. Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

@@

\section{DISCUSSION}

\subsection{Finding the Dominating Set with a Quantum Annealer}
\label{sec:discussion:qa}
\begin{figure}[b]
	\centering
		\includegraphics[width=0.5\columnwidth]{./figures/linear_crop.pdf}
	\caption{Linear graphs $G(n)$ used in this study. Nodes denote vertices of the graphs and lines are undirected edges.}
	\label{fig:linear}
\end{figure}

@@

\subsubsection{Modified annealing schedule}

@@

\subsection{Simulation Results}

@@

\subsection{Decoherence and Many-Body Localization}

@@

\section{METHODS}
\subsection{Simulation Details}
\subsubsection{The Lindblad Equation}
\label{sec:methods:lindblad}
Say what the simulator is solving

With anneal offsets

\subsubsection{Other Simulation Details}
Temperature
Initial wave function choices
RK tolerance

\subsubsection{Decoherence Time}
\label{sec:methods:deco}

\section{DATA AVAILABILITY}

@@

\section{ACKNOWLEDGEMENTS}

@@

\section{AUTHOR CONTRIBUTIONS}

@@

\section{ADDITIONAL INFORMATION}

@@

\bibliography{qilp}

>>>>

-\documentclass[prd,twocolumn,tightenlines,preprintnumbers,showpacs,superscriptaddress,notitlepage,nofootinbib,eqsecnum,floatfix,longbibliography]{revtex4}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{bm,bbm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{dsfont}

@@

%========================================================================================
\section{INTRODUCTION}
\label{sec:introduction}
%========================================================================================

@@

Formally, ILP is defined as the optimization of a linear function
{\color{blue}
\begin{align}
    f(x) = &\min\limits_{x}(\sum_i c_i x_i),\\
    \textrm{subject to} \quad & \sum_i A_{ai}x_i +b_a \leq 0 \\
    & x_i \geq 0,\\
    & x_i \in \mathbb{Z}
\end{align}
}
where $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations.

@@

%========================================================================================
\section{METHODS}
\label{sec:methods}
%========================================================================================

%----------------------------------------------------------------------------------------
\subsection{Integer Programming Mapping to QUBO}
\label{sec:methods:ilp}
%----------------------------------------------------------------------------------------

The inequality constraints can be rewritten as an equality by introducing slack variables $s$ such that the ILP becomes
{\color{blue}
\begin{align}
    f(x) = &\min(\sum_i c_i x_i),\\
    \textrm{subject to} \quad & \sum_i A_{a i}x_i + s_a + b_a = 0,\\
    & s_a \geq 0,\\
    & s_a \in \mathbb{Z},\\
    & x_i \geq 0,\\
    & x_i \in \mathbb{Z}.
\end{align}
}
Since quantum annealing solves the QUBO problem, we extend the proposed algorithm to solve integer quadratic optimization problems such that
{\color{blue}
\begin{align}
    f(x) = \min\limits_{x}(\sum_{ij} x_i d_{ij} x_j + \sum_i c_i x_i)
\end{align}
}
without the introduction of auxiliary qubits.

%----------------------------------------------------------------------------------------
\subsection{QUBO}
\label{sec:methods:qubo}
%----------------------------------------------------------------------------------------

We map integer variables $z_i$ to qubits under the following transformation~\cite{Chang:2018uoc}
\begin{align}
    z_i = & \sum_{r=0}^{R_i-1} 2^r \psi_{ri}
\label{eq:int_to_bin}
\end{align}
where $\psi_{ri} \in \{0, 1\}$ while the number of qubits used to represent the $i$-th integer variable is allowed to vary with $R_i$.
The transformation can in general be express as a rectangular matrix.
For example, a vector of two integer variables $z_0$ and $z_1$ represented by one and two qubits respectively is given as
\begin{align}
    \begin{pmatrix}
    z_0\\
    z_1
    \end{pmatrix}
    = &
    \begin{pmatrix}
    2^0 & 0 & 0\\
    0 & 2^0 & 2^1
    \end{pmatrix}
    \begin{pmatrix}
    \psi_{00}\\
    \psi_{01}\\
    \psi_{11}
    \end{pmatrix}
    \equiv T^z \begin{pmatrix}
    \psi_{00}\\
    \psi_{01}\\
    \psi_{11}
    \end{pmatrix}
\end{align}
The transformation can be reduced to a tensor product if $R_i$ is a constant for all $i$ such that
\begin{align}
    \mathcal{R} = & \begin{pmatrix} 2^0 & \dots & 2^{R-1}\end{pmatrix},\\
    \mathcal{Z} = & \begin{pmatrix} z_0 & \dots & z_{N-1}\end{pmatrix},\\
    |\mathds{1}| = & |\mathcal{Z}|,\\
    T^z = & \mathds{1}\otimes \mathcal{R}.
\end{align}

@@

%----------------------------------------------------------------------------------------
\subsection{Integer Programming Mapping to QUBO}
\label{sec:methods:ilp}
%----------------------------------------------------------------------------------------

The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
    \chi^2 = & \left[\Psi^x_{\mu} T^x_{\mu i}d_{ij} T^x_{j \nu}\Psi^x_\nu + c_i T^x_{i\mu} \Psi^x_\mu \right.
    \nonumber\\
    &\left.+ p (A_{a b} T^x_{b \mu} \Psi^x_{\mu} + T^s_{a \alpha} \Psi^s_\alpha + b_a)^2 \right],\\
    \Psi = & \begin{pmatrix}
    \Psi^x\\
    \Psi^s
    \end{pmatrix},
\end{align}
where $p$ is the strength of the penalty factor which must be set large enough such that the constraints are satisfied under minimization.
The objective function can be represented as a QUBO Hamiltonian
\begin{align}
    E = &
    \begin{pmatrix}
    \Psi^x & \Psi^s
    \end{pmatrix}
    \begin{pmatrix}
    Q_{xx} & Q_{xs}\\
    Q_{sx} & Q_{ss}
    \end{pmatrix}
    \begin{pmatrix}
    \Psi^x\\ \Psi^s
    \end{pmatrix} + pb^2,\\
    \equiv & \Psi^T Q \Psi + C,
    \label{eq:matrix_form}
\end{align}
where
\begin{align}
    Q_{xx} = & {T^{x}}^T \left[ d + p A^T A + p \mathrm{Diag}_{x} \left(A^T b + b^T A\right) \right] T^x \nonumber \\
    &+ \mathrm{Diag}_{\psi^x}(c) T^x,\\
    Q_{xs} = & Q_{sx}^T = p {T^{x}}^T A^T T^s,\\
    Q_{ss} = & p\left[ {T^{s}}^T T^s + \mathrm{Diag}_{\psi^s}\left( {T^{s}}^T b + b^T T^s\right) \right].
\end{align}
The function $\mathrm{Diag}_{z}(v)$ embeds a vector $v$ into a diagonal matrix in subspace $z$, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

%----------------------------------------------------------------------------------------
\subsection{Application to the Dominating Set Problem}
\label{sec:methods:mds}
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
\subsubsection{Minimum Dominating Set}
%----------------------------------------------------------------------------------------

The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
    f(x) = &\min\left(\sum_{\nu \in V} \psi^x_{\nu}\right),\\
    \textrm{subject to} \quad & \psi^x_{\nu} + \sum_{\mu \in \mathit{N}(\nu)} \psi^x_{\mu} \geq 1,\\
    & \psi^x_{\nu} \in \{0, 1\}
\end{align}
where the dimension of the dependent variable $x$ is the number of vertices $\overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\psi^x$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(\nu)$ is in $D$.
For each vertex in $V$ we introduce slack variables $s_{\nu} = T^s_{\nu \alpha} \psi^s_{\alpha}$ to encode the inequality constraint such that
\begin{align}
    f(x) = &\min(\sum_{\nu\in V} \psi^x_{\nu}),\\
    \textrm{subject to} \quad & \psi^x_{\nu} + J_{\nu \mu} \psi^x_{\mu}- T^s_{\nu \alpha} \psi^s_{\alpha}  - 1 = 0,\\
    & \psi_{\nu} \in \{0, 1\},\\
    & 0 \leq s_{\nu} \leq |\mathrm{N}(\nu)|\\
    & s_{\nu} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$ (symmetric and zero diagonal), the adjacency matrix for $G$.
The algorithm uses $N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits---$\overline{\overline{V}}$ qubits to encode the vertices and another $\sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits to embed the slack variables.
Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

@@

%----------------------------------------------------------------------------------------
\subsubsection{Modified annealing schedule}
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
\subsection{Simulation Results}
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
\subsection{Simulation Details}
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
\subsubsection{The Lindblad Equation}
\label{sec:methods:lindblad}
%----------------------------------------------------------------------------------------
Say what the simulator is solving

With anneal offsets
%----------------------------------------------------------------------------------------
\subsubsection{Other Simulation Details}
%----------------------------------------------------------------------------------------
Temperature
Initial wave function choices
RK tolerance

%----------------------------------------------------------------------------------------
\subsubsection{Decoherence Time}
\label{sec:methods:deco}
%----------------------------------------------------------------------------------------

%========================================================================================
\section{RESULTS \& DISCUSSION}
\label{sec:results}
%========================================================================================

%----------------------------------------------------------------------------------------
\subsection{Finding the Dominating Set with a Quantum Annealer}
\label{sec:discussion:qa}
%----------------------------------------------------------------------------------------

\begin{figure}[b]
    \centering
        \includegraphics[width=0.5\columnwidth]{./figures/linear_crop.pdf}
    \caption{Linear graphs $G(n)$ used in this study.
Nodes denote vertices of the graphs and lines are undirected edges.}
    \label{fig:linear}
\end{figure}

@@

%----------------------------------------------------------------------------------------
\subsubsection{Modified annealing schedule}
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
\subsection{Simulation Results}
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
\subsection{Decoherence and Many-Body Localization}
%----------------------------------------------------------------------------------------

@@

%========================================================================================
\section{DATA AVAILABILITY}
%========================================================================================

@@

%========================================================================================
\section{ACKNOWLEDGEMENTS}
%========================================================================================

@@

%========================================================================================
\section{AUTHOR CONTRIBUTIONS}
%========================================================================================

@@

%========================================================================================
\section{ADDITIONAL INFORMATION}
%========================================================================================

@@

%========================================================================================
\bibliography{qilp}
%========================================================================================
